---
title: "PA_Template1"
author: "Zack Gouttel"
date: "15/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Assignment Report

This is an R Markdown made docment for the Assignment report of Reproducible Research's 1st Course Project

# 1.Reading the data into R
 I used read.csv:
```{r}
activity <- read.csv("~/GitHub/RR_CP1/repdata_data_activity/activity.csv")
```


# 2.a Calculating the total number of steps per day and plotting it

Did this through the aggregate function
```{r}
sum_steps<- aggregate(activity[, 1], list(activity$date), sum, na.rm=TRUE)
```




```{r steps, echo=FALSE}
 hist(sum_steps[,2], main="Total Step Count Histogram")
```

# 2.b Mean and Median

The mean and median are as follows in the summary function

```{r}
 summary(sum_steps[,2])
```


# 3. Average Daily Activity Pattern
 This was done in two steps
 
## First one
 
 Calculating the average daily pattern for each 5 minute interval each day through the aggregate function
 
```{r,echo=FALSE}
   mean_steps<-aggregate(activity[,1], list(activity$date,activity$interval),
                          mean,
                      na.rm=TRUE)
```

## Second One

 Plotting
```{r Daily Pattern, echo=FALSE}
 plot(x = mean_steps[,2],y=mean_steps[,3],xlab="interval",
     ylab="Average step count", main = "Average daily pattern",type = "l")
    abline(h=max(mean_steps[,3],na.rm = TRUE))   
```
 
 
 Viewing the Summary of the average daily pattern returns that the max value is 806, equivalent to the interval 615
 
```{r}
 summary(mean_steps[,3])
```
# 4.Imputing NAs:

  We used a simple manner to do so by replacing the 2304 NAs by the mean of activity$steps, then we created a new NA clean Dataframe called dfnna
  
```{r,echo=FALSE}
 summary(activity)                                                                #it returns that activity contains 2304 missing value
                                                                         
     dfnna<-activity                                                                 #here i chose to replace the NAs in the whole dataset
                                                                  
     dfnna[is.na(activity)]<-mean(activity$steps,na.rm = TRUE)                       #by the mean of activity$steps
     
```
  
## 4.1 Plotting the new total average number of steps per day and gathering the new mean and median
  This was done in the same manner as the original dataset
  
```{r,echo=FALSE}
 sum_steps_nna<- aggregate(dfnna[, 1], list(dfnna[,2]), sum)                      
     
```  
  
  then the plot: 
  
```{r Steps_no_na, echo=FALSE}
hist(sum_steps_nna[,2], main="Total Step Count Histogram -NO NAs-")   
```  
  
  The SUMMARY reports that the median and the mean are now equal 

```{r}
 summary(sum_steps_nna[,2])
```

# 5. Difference in Average Daily Pattern between Weekdays and Weekends

 creating a new factor for weekend and weekday through the weekdays function
 
```{r}
 r<-weekdays(as.Date(dfnna$date))  
we<-ifelse(r %in% c("samedi", "dimanche"), "weekend", "weekday") 
dfnna$weekend_weekday<-we 
mean_steps_nna<-aggregate(dfnna[,1], list(dfnna$weekend_weekday,dfnna$interval),
                         mean,
                         na.rm=TRUE)

```
  Plotting the Panel Plot:
  
```{r Panel_plot, echo=FALSE}
 qplot(mean_steps_nna[,2],mean_steps_nna[,3],data=mean_steps_nna,facets = .~Group.1, geom="smooth")    
```    

We can observe that average daily pattern is relatively higher in weekends than weekdays, which is interesting

# This Concludes my Report for the Course Project.